The formal verification of the ctm approach to forcing \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2022-06-07 Independence_CH/related.tex
Comments for further edits, 2nd part
Some points for Isar.

This faithfulness to set-theoretical practice does not come for
free. Recursive constructions and inductive definitions are far easier
to perform in the Calculus of Inductive Constructions on which Lean
(and Coq) are based, and in Isabelle/ZF are (still) rather
cumbersome. Also, a typed discipline provides aid to write succintly
and many assumptions are satisfied by mere notation.

%% Comment on ZFC_in_HOL

>>>>

This faithfulness to set-theoretical practice does not come for
free. Recursive constructions and inductive definitions are far easier
to perform in the Calculus of Inductive Constructions on which Lean
(and Coq) are based, and in Isabelle/ZF are rather cumbersome. Also, a
typed discipline provides aid to write succintly and many assumptions
are satisfied by mere notation. To be clear, those benefits come from
doing set theory in a non set-theoretical meta-language. On the other
hand, Isar proofs, as the one shown in
Sec.~\ref{sec:sample-formal-proof}, are easier to write and understand
than the language of tactics of Lean.
%% Comment on ZFC_in_HOL and other developments using it (mostly by
%% Paulson: Ramsey theory, (ordinal) partition theorems)
